Mining Scenario/Transaction Pairs HW/SW Co-verification Using Symbolic Execution

### 2017-04-03 paper.tex
circuit to software

\textbf{Co-specification Model} We verify that the transmitted data is the same as the 
received data in the {\em loopback} mode.  To do so, we annotate the FW model 
with an assertion $assert()$, as shown in Figure~\ref{firmware}.  Note that the 
array containers $rx\_b$ and $tx\_b$ stores the data obtained through the $inb$
and $outb$ function calls which directly read from and write to the UART IP ports
respectively. 
\[
  for(i=0; i<=N; i++) 
   \quad assert(rx\_b[i] == tx\_b[i]);
\]
%
We formally describe the notion of \emph{transaction} and \emph{scenario} with the
help of our working example. \\
%
\begin{figure}[t]
\scriptsize
\begin{tabular}{l|l}
\hline
Wishbone Interface & Main Module \\
\hline
\begin{lstlisting}[mathescape=true,language=C]
typedef unsigned char u8;
unsigned char inb 
 (unsigned long port) {
  return wb_read(port);
}
void outb (u8 value, 
 unsigned long port) {
  wb_write(port, value);
}
void wb_reset(void) {
 rst_i = 1;
 set_inputs();
 next_timeframe();
 rst_i = 0;
 stb_i = 0; 
 cyc_i = 0;
}
void wb_idle() {
 set_inputs();
 next_timeframe();
}
void wb_write(_u32 addr, 
     _u8 b) {
 adr_i = addr;
 dat_i = b;
 we_i = 1;
 cyc_i = 1;
 stb_i = 1;
 set_inputs();
 next_timeframe();
 we_i = 0;
 cyc_i = 0;
 stb_i = 0;
}
\end{lstlisting}
&
\begin{lstlisting}[mathescape=true,language=C]
int main() {
wb_reset();
wb_idle();
// Configure the uart
outb (0x13, UART_MC);  
outb (0x80, UART_CM3);
outb (0x00, UART_CM2);
outb (0x00, UART_CM1);
outb (0x00, UART_CR);  
outb (0x03, UART_IE); 
char tx_b[] = "Hello world";
_u8 status = 0;
char rx_b[100];
int i=0,c=0,d=0;
// data transfer in loopback
for (i=0; i<1990; i++){
if (irq_o){
 status=inb(UART_IS)&0x0c;
 if(istatus==0x0c){
 //it was a tx_empty interrupt
 outb(*(tx_b+c),UART_TR); c++;
 }
 else{ //status==0x04
 //it was an rx_data interrupt
  rx_b[d] = inb(UART_TR); d++;
 }
}else {
 // no interrupt. 
 wb_idle();
 wb_idle();
 }
}
// property
for(i=0; i<=10; i++)
 assert(rx_b[i] == tx_b[i]);
}
\end{lstlisting}
\\
\hline
\end{tabular}
\caption{Firmware model of UART IP}
\label{firmware} 
\end{figure}
%

@@

Figure~\ref{fig:interleaving2} shows a fragment of $\mathcal{HT}$
corresponding to the top module of UART IP.  Let us consider a possible
execution of $\mathcal{HT}$, $(init \rightarrow t_1 \rightarrow t_4
\rightarrow t_6 \rightarrow comp)$.  At $init$, the FW sets the HW input
signal, $(rst\_i==0)$, based on which the execution follows the edge to
$t_1$.  At $t_1$, the predicate, $cs$, is computed based on HW input
signals, $cyc\_i$, $stb\_i$, $adr\_i$ and $we\_i$.  Assuming that $cs\_i
\&\& we\_i$ holds true, the execution proceeds to step $t_4$ which computes
predicates based on the values in $adr\_i$.  The next step is determined by
the values of predicates, $'UART\_MC, 'UART\_IE, 'UART\_CT$.  Assuming that
$'UART\_MC$ holds true, the execution proceeds to $t_6$.  During the
execution of $t_6$, three parallel state update $(dtr\_no, rts\_no,
loopback)$ and two predicate evaluation ($\beta_1,\beta_2)$ is performed
based on the updates to state elements.  After $t_6$, the execution of
$\mathcal{HT}$ leads to $comp$ and the transaction finishes its execution. 
Note that the dotted edges in $\mathcal{HT}$ denotes that eventually the
execution reaches $comp$.  The update of state elements are shown in
rectangular boxes, whereas the evaluation of boolean predicates are shown
inside the rounded boxes.  The blue arrows point to the state updates or
predicate evaluation at a particular~$\varphi$. \\
%
\begin{figure}[t]
\scalebox{.55}{\import{figures/}{fw-hw-interleaving.pspdftex}}
\caption{Some HW Transactions for UART
\label{fig:interleaving2}}
\end{figure}
%
\textbf{Scenario/Transaction Pairs}
%
A scenario, $\mathcal{S}$, restricts the interaction between $\mathcal{HT}$
and $\mathcal{FT}$.  Similar to service-function based transaction in
TLM~\cite{hvc}, we view $\mathcal{FT}$ and $\mathcal{HT}$ as a set of
service in response to its interacting $\mathcal{HT}$ or $\mathcal{FT}$,
respectively.  A {\em Scenario/Transaction} pair is a set of valid
interactions between $\mathcal{HT}$ and $\mathcal{FT}$ for a
particular~$\mathcal{S}$.  For example, an UART IP contains a {\em
transmitter} and a {\em receiver} module.  Let us assume that a scenario,
$\mathcal{S}$, configure a UART in {\em transmission} mode.  Then the
interaction between $\mathcal{FT}$ and $\mathcal{HT}$ is constrained to
contain only those interactions that are relevant to the {\em transmitter}
module.  By contrast, the interactions with respect to the {\em receiver}
module is irrelevant for $\mathcal{S}$.  However, when $\mathcal{S}$
configures a UART in {\em loopback} mode, then the interaction between
$\mathcal{FT}$ and $\mathcal{HT}$ must contain transactions in {\em
transmitter} as well as {\em receiver} module.
%
\Omit{
\textbf{Problem Statement}
Given a FW model in C, HW models in Verilog RTL, a property $\psi$ and 
a set of scenarios, the goal is to check whether $\psi$ holds under all 
scenarios in the combined FW-HW model. 
}
%
\section{Proposed Methodology}

@@

to $Out$, $Seq$ and $Comb$. \\ 

@@

\textbf{Step 2: Constructing Co-verification Model}
An unified co-verification model is constructed by sequentializing 
HW and its interacting FW threads, thereby generating a sequential 
single-threaded co-verification model in C.  The motivation behind
sequentialization is based on our observation that practical industrial
co-designs exhibit producer-consumer interaction behavior.  We observe 
that in practise, a FW thread is mostly independent of the HW thread they interact 
with~\cite{hvc,codes14}.  This behavior is also prevelant in industrial co-designs, 
such as IBM Text Accelerator co-design~\cite{polig2014micro,polig2014fpl}
and IBM coherent FFT coprocessor~\cite{giefers2015accelerating}.
Specifically, a FW thread is only responsible for configuring the 
HW transactions by writing to memory mapped registers or polling 
the interrupt status register for data transmission or receiving 
incoming data packets. This interaction pattern exhibits producer-consumer 
relationship, which is also evident in case of our UART example in 
Figure~\ref{fig:interleaving1}.  
We formally define a single-threaded sequential unified co-verification model, 
denoted by $\mathcal{M}_{seq}$. 
  \mathcal{M}_{seq} & {:}{:}{=} & \mathit{Procedure} \\
\mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
\mathit{Statement} & {:}{:}{=} & CStatement \mid \mathit{moduleName(Var_1,\dots,Var_n)} \\
\mathit{CStatement} & {:}{:}{=} & x {:}{=} exp \mid \mathit{ITE}(b, s_1,s_2) \mid \mathit{s_1;s_2} \mid loop\{s\} \\ 
Consider sets of expressions $Exp$ and Boolean expressions $BExp$ 
over variables $Var$ of $\mathcal{M}_{seq}$ such that $(exp \in Exp)$ and 
$(b \in BExp)$. A $\mathit{CStatement}$ is an assignment,
conditional, sequential concatenation or a loop.
$\mathit{moduleName(Var_1,\dots,Var_n)}$ invokes the top level 
procedure of the $\mathcal{SN}$. Recall that $\mathcal{SN}$ is a
software netlist representation of HW RTL models.
The interface variables $\{Var\_1,\dots,Var\_n\}$ 
(input, output and inout ports) are declared inside the 
FW to communicate with $\mathcal{SN}$. \\ 
%\textbf{Producer-consumer Interaction:}
A communicating pair of executions of 
$\mathcal{FT}$ and $\mathcal{HT}$ is an {\em interaction}, which 
takes place through shared variables (or shared registers)
only.  An interaction is said to exhibit {\em producer-consumer}
interaction pattern if 1) the set of variables $V$ written by 
the execution of $\mathcal{FT}$ and and the set of variables 
$V'$ written by the execution of $\mathcal{HT}$ are disjoint 
($V \cap V' = \emptyset$), (2) and the executions 
of $\mathcal{FT}$ and $\mathcal{HT}$ consumes values 
in $V'$ and $V$, respectively.

>>>>

\textbf{Co-specification Model} We verify that the transmitted data is the same as the 
received data in the {\em loopback} mode.  To do so, we annotate the FW model 
with an assertion $assert()$, as shown in Figure~\ref{firmware}.  Note that the 
array containers $rx\_b$ and $tx\_b$ stores the data obtained through the $inb$
and $outb$ function calls which directly read from and write to the UART IP ports
respectively. \rmcmt{more on co-specification PSL}. 
\[
  for(i=0; i<=N; i++) 
   \quad assert(rx\_b[i] == tx\_b[i]);
\]
%
We formally describe the notion of \emph{transaction} and \emph{scenario} with the
help of our working example. \\
%
\begin{figure}[t]
\scriptsize
\begin{tabular}{l|l}
\hline
Wishbone Interface & Main Module \\
\hline
\begin{lstlisting}[mathescape=true,language=C]
typedef unsigned char u8;
unsigned char inb 
 (unsigned long port) {
  return wb_read(port);
}
void outb (u8 value, 
 unsigned long port) {
  wb_write(port, value);
}
void wb_reset(void) {
 rst_i = 1;
 set_inputs();
 next_timeframe();
 rst_i = 0;
 stb_i = 0; 
 cyc_i = 0;
}
void wb_idle() {
 set_inputs();
 next_timeframe();
}
void wb_write(_u32 addr, 
     _u8 b) {
 adr_i = addr;
 dat_i = b;
 we_i = 1;
 cyc_i = 1;
 stb_i = 1;
 set_inputs();
 next_timeframe();
 we_i = 0;
 cyc_i = 0;
 stb_i = 0;
}
\end{lstlisting}
&
\begin{lstlisting}[mathescape=true,language=C]
int main() {
wb_reset();
wb_idle();
// Configure the uart
outb (0x13, UART_MC);  
outb (0x80, UART_CM3);
outb (0x00, UART_CM2);
outb (0x00, UART_CM1);
outb (0x00, UART_CR);  
outb (0x03, UART_IE); 
char tx_b[] = "Hello world";
_u8 status = 0;
char rx_b[100];
int i=0,c=0,d=0;
// data transfer in loopback
for (i=0; i<1990; i++){
if (irq_o){
 status=inb(UART_IS)&0x0c;
 if(istatus==0x0c){
 //it was a tx_empty interrupt
 outb(*(tx_b+c),UART_TR); c++;
 }
 else{ //status==0x04
 //it was an rx_data interrupt
  rx_b[d] = inb(UART_TR); d++;
 }
}else {
 // no interrupt. 
 wb_idle();
 wb_idle();
 }
}
// property
for(i=0; i<=10; i++)
 assert(rx_b[i] == tx_b[i]);
}
\end{lstlisting}
\\
\hline
\end{tabular}
\caption{Firmware model of UART IP}
\label{firmware} 
\end{figure}
%

@@

Figure~\ref{fig:interleaving2} shows a fragment of $\mathcal{HT}$
corresponding to the top module of UART IP.  Let us consider a possible
execution of $\mathcal{HT}$, $(init \rightarrow t_1 \rightarrow t_4
\rightarrow t_6 \rightarrow comp)$.  At $init$, the FW sets the HW input
signal, $(rst\_i==0)$, based on which the execution follows the edge to
$t_1$.  At $t_1$, the predicate, $cs$, is computed based on HW input
signals, $cyc\_i$, $stb\_i$, $adr\_i$ and $we\_i$.  Assuming that $cs\_i
\&\& we\_i$ holds true, the execution proceeds to step $t_4$ which computes
predicates based on the values in $adr\_i$.  The next step is determined by
the values of predicates, $'UART\_MC, 'UART\_IE, 'UART\_CT$.  Assuming that
$'UART\_MC$ holds true, the execution proceeds to $t_6$.  During the
execution of $t_6$, three parallel state update $(dtr\_no, rts\_no,
loopback)$ and two predicate evaluation ($\beta_1,\beta_2)$ is performed
based on the updates to state elements.  After $t_6$, the execution of
$\mathcal{HT}$ leads to $comp$ and the transaction finishes its execution. 
Note that the dotted edges in $\mathcal{HT}$ denotes that eventually the
execution reaches $comp$.  The update of state elements are shown in
rectangular boxes, whereas the evaluation of boolean predicates are shown
inside the rounded boxes.  The blue arrows point to the state updates or
predicate evaluation at a particular~$\varphi$. \\
\scalebox{.55}{\import{figures/}{fw-hw-interleaving.pspdftex}}
\caption{Some HW Transactions for UART
\label{fig:interleaving2}}
\textbf{Scenario/Transaction Pairs}
A scenario, $\mathcal{S}$, restricts the interaction between $\mathcal{HT}$
and $\mathcal{FT}$.  Similar to service-function based transaction in
TLM~\cite{hvc}, we view $\mathcal{FT}$ and $\mathcal{HT}$ as a set of
service in response to its interacting $\mathcal{HT}$ or $\mathcal{FT}$,
respectively.  A {\em Scenario/Transaction} pair is a set of valid
interactions between $\mathcal{HT}$ and $\mathcal{FT}$ for a
particular~$\mathcal{S}$.  For example, an UART IP contains a {\em
transmitter} and a {\em receiver} module.  Let us assume that a scenario,
$\mathcal{S}$, configure a UART in {\em transmission} mode.  Then the
interaction between $\mathcal{FT}$ and $\mathcal{HT}$ is constrained to
contain only those interactions that are relevant to the {\em transmitter}
module.  By contrast, the interactions with respect to the {\em receiver}
module is irrelevant for $\mathcal{S}$.  However, when $\mathcal{S}$
configures a UART in {\em loopback} mode, then the interaction between
$\mathcal{FT}$ and $\mathcal{HT}$ must contain transactions in {\em
transmitter} as well as {\em receiver} module.
\section{Proposed Methodology}

@@

to $Out$, $Seq$ and $Comb$. Figure~\ref{ex1} shows an example of 
translating a Verilog RTL circuit with an always block and continuous
assignment statement into an equivalent C program.

@@

\begin{figure*}[t]
\scriptsize  
\centering
\begin{tabular}{l|l|l}
\hline
 Verilog & Dataflow Graph & Software netlist \\
\hline
\begin{lstlisting}[mathescape=true,language=Verilog]
module top(clk, a);
input clk, a;
reg b, d, e;
wire c = e ? 1'b0:d;
wire cond = a;
always @(posedge clk) 
 begin
  b<=a;
  if(cond && b)
   e<=b;
  else 
   e<=0;
  d<=c;
 end
endmodule
\end{lstlisting}
&
\begin{minipage}{5.0cm}
\centering
\scalebox{.5}{\import{figures/}{ckt.pspdftex}}
\end{minipage}
&
\begin{lstlisting}[mathescape=true,language=C]
struct state_elements_top {
 unsigned int b, d, e, c;
};
struct state_elements_top  u1;
void top(unsigned int clk, 
   unsigned int a)
{
  _Bool b_old,d_old,e_old;
  _Bool cond;

  b_old = u1.b;
  d_old = u1.d;
  e_old = u1.e;

  cond = a;
  u1.c = (u1.e) ? 0 : u1.d;

 u1.b = a;
 if(cond && b_old)
    u1.e = b_old;
  else
    u1.e = 0;
  u1.d = u1.c;  
}
\end{lstlisting}
\\
\hline
\end{tabular}
\caption{Circuit to Software}
\label{ex1}
\end{figure*}
%
\textbf{Step 2: Constructing Co-verification Model}
An unified co-verification model is constructed by sequentializing 
HW and its interacting FW threads, thereby generating a sequential 
single-threaded co-verification model in C.  The motivation behind
sequentialization is based on our observation that many practical industrial
co-designs exhibit producer-consumer interaction behavior, such as IBM 
Text Accelerator co-design~\cite{polig2014micro,polig2014fpl}, 
IBM coherent FFT coprocessor~\cite{giefers2015accelerating}, RockBox 
Media Player~\cite{hvc}, linux device driver interacting with QEMU emulated
hardware code~\cite{codes14} and others.  We observe that in practise, a FW 
thread is mostly independent of the HW thread they interact with~\cite{hvc,codes14}.  
Specifically, a FW thread is only responsible for configuring the 
HW transactions by writing to memory mapped registers or polling 
the interrupt status register for data transmission or receiving 
incoming data packets. This interaction pattern exhibits producer-consumer 
relationship, which is also evident in case of our UART example in 
Figure~\ref{fig:interleaving1}.  
%
We now formally define a single-threaded sequential co-verification model, 
denoted by $\mathcal{M}_{seq}$, obtained by sequentializing the HW and FW 
models.
%
\[ 
\begin{array}[t]{@{}lll}
  \mathcal{M}_{seq} & {:}{:}{=} & \mathit{Procedure} \\
\mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
\mathit{Statement} & {:}{:}{=} & CStatement \mid \mathit{moduleName(Var_1,\dots,Var_n)} \\
\mathit{CStatement} & {:}{:}{=} & x {:}{=} exp \mid \mathit{ITE}(b, s_1,s_2) \mid \mathit{s_1;s_2} \mid loop\{s\} \\ 
\end{array}
\]
%
Consider sets of expressions $Exp$ and Boolean expressions $BExp$ 
over variables $Var$ of $\mathcal{M}_{seq}$ such that $(exp \in Exp)$ and 
$(b \in BExp)$. A $\mathit{CStatement}$ is an assignment,
conditional, sequential concatenation or a loop.
$\mathit{moduleName(Var_1,\dots,Var_n)}$ invokes the top level 
procedure of the $\mathcal{SN}$. Recall that $\mathcal{SN}$ is a
software netlist representation of HW RTL models.
The interface variables $\{Var\_1,\dots,Var\_n\}$ 
(input, output and inout ports) are declared inside the 
FW to communicate with $\mathcal{SN}$. \\ 
%
%\textbf{Producer-consumer Interaction:}
%
A communicating pair of executions of 
$\mathcal{FT}$ and $\mathcal{HT}$ is an {\em interaction}, which 
takes place through shared variables (or shared registers)
only.  An interaction is said to exhibit {\em producer-consumer}
interaction pattern if 1) the set of variables $V$ written by 
the execution of $\mathcal{FT}$ and and the set of variables 
$V'$ written by the execution of $\mathcal{HT}$ are disjoint 
($V \cap V' = \emptyset$), (2) and the executions 
of $\mathcal{FT}$ and $\mathcal{HT}$ consumes values 
in $V'$ and $V$, respectively.
